---
title: "1. Introduction to R "
author: "Sonali Arora"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
vignette: >
  % \VignetteIndexEntry{1. Introduction to Bioconductor}
  % \VignetteEngine{knitr::rmarkdown}
---
```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=100, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")))
```

Author: Sonali Arora (<a
  href="mailto:sarora@fredhutch.org">sarora@fredhutch.org</a>)<br/ >
Date: 20-22 July, 2015<br />

The material in this course requires R version 3.2.1 and Bioconductor
version 3.2

## R 

- http://r-project.org
- Open-source, statistical programming language; widely used in academia, finance, pharma, . . .
- Core language, ‘base’ and > 4000 contributed packages
- Interactive sessions, scripts, packages

## Data types in R 

- Vectors 
  – logical, integer, numeric, character, . . .  
  - list() – contains other vectors (recursive)
  - factor(), NA – statistical concepts  
  - Can be named – c(Seattle=1, Portland=2)  
- matrix(), array() – a vector with a ‘dim’ attribute.  
- data.frame()  
  – like spreadsheets; list of equal length vectors.
  - Homogenous types within a column, heterogenous types across columns.  
- Other classes 
  – more complicated arrangement of vectors.
  - Examples  
     - the value returned by lm(); 
     - the DNAStringSet class used to hold DNA sequences.
  - plain, ‘accessor’, ‘generic’, and ‘method’ functions
- Packages 
  – base, recommended, contributed.






## `sessionInfo()`

```{r sessionInfo}
sessionInfo()
```

