---
title: "1. Introduction to Bioconductor "
author: "Sonali Arora"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
vignette: >
  % \VignetteIndexEntry{1. Introduction to Bioconductor}
  % \VignetteEngine{knitr::rmarkdown}
---
```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=100, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")))
```
Author: Sonali Arora (<a
  href="mailto:sarora@fredhutch.org">sarora@fredhutch.org</a>)<br/ >
Date: 20-22 July, 2015<br />

The material in this course requires R version 3.2.1 and Bioconductor
version 3.2


## What is Bioconductor 

Analysis and comprehension of high-throughput genomic data

- Statistical analysis: large data, technological artifacts, designed
  experiments; rigorous
- Comprehension: biological context, visualization, reproducibility
- High-throughput
    + Sequencing: RNASeq, ChIPSeq, variants, copy number, ...
    + Microarrays: expression, SNP, ...
    + Flow cytometry, proteomics, images, ...

Packages, vignettes, work flows

- 'Release' (every six months) and 'devel' branches
- Choose from 1045 packages.
     + Discover packages via 
     [biocViews](http://bioconductor.org/packages/devel/BiocViews.html#___Software)
     + Each Package has a Package 'landing page' with Title, 
       author / maintainer, short description, citation, 
       installation instructions, shields, documentation, download statistics 
     + All user-visible functions have help pages, most with runnable examples
     + 'Vignettes' an important feature in Bioconductor -- narrative
       documents illustrating how to use the package, with integrated code

Helpful Links  

- Documentation
[Previous Course Material](http://bioconductor.org/help/course-materials/),
[Workflows](http://bioconductor.org/help/workflows/),
[Videos](https://www.youtube.com/user/bioconductor),
[Developers](http://bioconductor.org/developers/) 
- Ask a question 
[Support site](https://support.bioconductor.org) 
- Connect with us 
[Twitter](https://twitter.com/Bioconductor), 
[Newsletter](http://bioconductor.org/help/newsletters/2015_July/)


## Overall Workflow & How Bioconductor fits in 

A typical workflow consists of the following steps.  
1. Experimental design  
2. Wet-lab preparation  
3. High-throughput sequencing  
       + Output: FASTQ files of reads and their quality scores  
4. Alignment
       + Many different aligners, some specialized for different purposes  
       + Output: BAM files of aligned reads  
5. Summary  
       + e.g., _count_ of reads overlapping regions of interest (e.g., genes)  
6. Statistical analysis  
7. Comprehension  

## `sessionInfo()`

```{r sessionInfo}
sessionInfo()
```










  
